Predictive Quantization for MIMO-OFDM SVD Precoders using Reservoir Computing Framework

### 2019-04-15 paper/main.tex
More changes

% Elaborate on each point (what people have done before) lecunnas
This challenge has been overcome in past work via two inherent properties of the precoders. 
Firstly, the precoders are not actually $2\times N_T\times N_R$ dimensional entities, and these actually have an underlying manifold structure which allows one to work with lower dimensions if one performs operations over the manifold directly.
Secondly, these precoders are correlated nicely along both time and frequency. 
Capturing temporal correlations allow for predictive quantization algorithms, which allow improvement of the quantization error with time.
Exploiting frequency correlations allow for various interpolation algorithms, which in turn reduce the feedback overhead in a single OFDM frame, by feeding back only for certain subcarriers, and interpolating over the others.
Although the underlying manifold structure allows one to work with effective lower dimensions, most of the naive linear algorithms to predict and interpolate don't work well here, owing to the inherent non-linear structure of the manifold.
Hence, combining these two properties, viz. the underlying manifold structure and temporal-frequency correlations, require generalization of such linear algorithms to the manifold under consideration, by exploiting the differential geometry of the manifold. Relevant past work in this direction has been \cite{Gupt1905:Predictive,6891198,6545375,5671092,Li2016,khaled2005quantized,krishnamachari2013geometry}.

% What is the way ahead
The manifold based approaches have provided significant performance benefits, given that they work with effective lower dimensions.
However, these approaches are difficult to implement and require some familiarity with manifold geometry for a complete grasp.
In this paper, we propose a reservoir computing based predictive quantizer to exploit the temporal correlations effectively.
The proposed scheme offers a cleaner solution to the predictive quantization problem, which brings about both ease of implementability, and improved performance as well, when compared to \cite{6891198}, which proposed a manifold based method of predictive quantization.

% How to we do this 
Reservoir computing is a computational framework designed for sequential data processing. 
It's design takes inspiration from several frameworks of recurrent neural networks \cite{jaeger2004harnessing} including Liquid State Machines \cite{jalalvand2015real}. 
A reservoir computing system maps the inputs to spatio-temporal patterns in higher dimensional dynamical reservoir state. 
The pattern analysis over these spatio-temporal patterns is performed by an output coupler which maps the higher dimensional reservoir state back to the output dimension, generally having lower dimensionality than reservoir state. 
% A important characteristic of reservoir computing is that while training, the input weights and the weights within the reservoir are not kept fixed and the weights of the output coupler are updated. 
The major advantage of reservoir computing system is that it’s extremely easy to train in comparison to its counterparts like Deep Neural Networks. 
The input features can be efficiently read out by a simple learning algorithm because the reservoir non-linearly transforms sequential inputs into the higher dimensional reservoir state. 
Reservoir computing has been successfully used to solve problems like spoken digit recognition \cite{verstraeten2005isolated}, handwritten digit image recognition \cite{jalalvand2015real} and climate prediction \cite{pathak2017using}.  
% Owing to the versatility of reservoir computing, it can deal with any sequential data in principle and hence further expansion of its application fields is widely expected.

Reservoir computing has also been applied to solve problems in the domain of wireless communications \cite{jaeger2004harnessing,mosleh2017brain,shafin2018realizing}. 
\cite{jaeger2004harnessing} used reservoir computing framework to harness channel non-linearities in order for smarter channel equalization.
\cite{mosleh2017brain,shafin2018realizing} utilized the same for OFDM symbol detection while tapping in channel, as well as power amplifier non-linearities.
In our work, we utilize the reservoir computing framework to discover the underlying non-linear relations between previously observed precoding matrices, in order to predict the next precoder in series. Upon vectorizing the precoding matrices, this becomes analogous to time series prediction algorithms, which the reservoir computing framework has been known to work really well for \cite{mosleh2017brain}.

>>>>

% Elaborate on each point (what people have done before) lecunnas
Past work has used two inherent properties of the precoders to enable effective low-rate precoder feedback. 
First, the precoders are not actually $2\times N_T\times N_R$ dimensional entities, and these have an underlying manifold structure that allows one to work with lower dimensions if one performs operations over the manifold directly.
Second, these precoders are correlated in both time and frequency. 
Capturing temporal correlations allows the use off predictive quantization algorithms that improvement of the quantization error with time by effectively utilizing all information.
Exploiting frequency correlations permits the use of interpolation algorithms, which in turn reduce the feedback overhead in a single OFDM frame, by feeding back CSI information only for certain subcarriers, and interpolating over the others.
Despite the existence of the underlying manifold structure that possesses effective lower dimensions, most of the na{\"i}ve algorithms that predict and interpolate are limited in their effectiveness, owing to the inherent non-linear structure of the manifold.
Thus, combining these two properties, viz. the underlying manifold structure and temporal-frequency correlations, requires a generalization of such linear algorithms to the manifold under consideration, by exploiting the differential geometry of the manifold. Relevant past work in this direction has been \cite{Gupt1905:Predictive,6891198,6545375,5671092,Li2016,khaled2005quantized,krishnamachari2013geometry}.

% What is the way ahead
Manifold based approaches have provided significant performance benefits, given that they work with effective lower dimensions.
However, these approaches are difficult to implement and require some familiarity with manifold geometry for a complete grasp.
In this paper, we propose a reservoir computing based predictive quantizer to exploit the temporal correlations effectively.
The proposed scheme offers a cleaner solution to the predictive quantization problem, which brings about both ease of implementability, and improved performance as well, when compared to \cite{6891198}, which proposed a manifold based method of predictive quantization.

% How to we do this 
Reservoir computing is a computational framework designed for sequential data processing. 
Its design is inspired from several frameworks of recurrent neural networks \cite{jaeger2004harnessing}, including Liquid State Machines \cite{jalalvand2015real}. 
A reservoir computing system maps the inputs to spatio-temporal patterns in a higher dimensional dynamical reservoir state. 
The pattern analysis over these spatio-temporal patterns is performed by an output coupler that maps the higher dimensional reservoir state back to the output dimension that generally possesses lower dimensionality than the reservoir state. 
% A important characteristic of reservoir computing is that while training, the input weights and the weights within the reservoir are not kept fixed and the weights of the output coupler are updated. 
The key advantage of a reservoir computing based system is that it is extremely easy to train in comparison to its counterparts like Deep Neural Networks. CITATION NEEDED
The input features can be efficiently learnt using a simple learning algorithm, since the reservoir non-linearly transforms sequential inputs into a higher dimensional reservoir state. 
Reservoir computing has been successfully used to solve problems like spoken digit recognition \cite{verstraeten2005isolated}, handwritten digit image recognition \cite{jalalvand2015real} and climate prediction \cite{pathak2017using}.  
% Owing to the versatility of reservoir computing, it can deal with any sequential data in principle and hence further expansion of its application fields is widely expected.

Reservoir computing has also been applied to solve problems in the domain of wireless communications \cite{jaeger2004harnessing,mosleh2017brain,shafin2018realizing}. 
\cite{jaeger2004harnessing} used reservoir computing framework to harness channel non-linearities in order for smarter channel equalization.
\cite{mosleh2017brain,shafin2018realizing} utilized this technique for OFDM symbol detection while accounting for channel and power amplifier non-linearities.
In our work, we utilize the reservoir computing framework to discover the underlying nonlinear relations between previously observed precoding matrices, in order to predict the next precoder in series.
Upon vectorizing the precoding matrices, this becomes analogous to time series prediction algorithms, for which the reservoir computing framework has been known to work effectively \cite{mosleh2017brain}.
